%-------------------------
% Resume in Latex
% Author : Jake Gutierrez
% Based off of: https://github.com/sb2nov/resume
% License : MIT
%------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\input{glyphtounicode}

%----------FONT OPTIONS----------
% sans-serif
% \usepackage[sfdefault]{FiraSans}
% \usepackage[sfdefault]{roboto}
% \usepackage[sfdefault]{noto-sans}
% \usepackage[default]{sourcesanspro}
%
% serif
% \usepackage{CormorantGaramond}
% \usepackage{charter}

\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

% Ensure that generated pdf is machine readable / ATS parsable
\pdfgentounicode=1

%-------------------------
% Custom commands

\newcommand{\resumeItem}[1]{
  \item\small{
    {#1 \vspace{-2pt}}
  }
}

% <<< UPDATED: added positive vspaces before & after each subheading >>>
\newcommand{\resumeSubheading}[4]{
  \vspace{4pt}  % space before heading
  \item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{2pt}  % space after heading
}

\newcommand{\resumeSubSubheading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small#1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeProjectHeading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \small#1 & #2 \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubItem}[1]{\resumeItem{#1}\vspace{-4pt}}

\renewcommand\labelitemii{$\vcenter{\hbox{\tiny$\bullet$}}$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.15in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
%%%%%%  RESUME STARTS HERE  %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%----------HEADING----------
\begin{center}
    \textbf{\Huge \scshape Hamza Khan} \\ \vspace{1pt}
    \small +92-347-8479617$|$ 
    \href{mailto:muhammadhamzakhanmarwat@gmail.com}{\underline{muhammadhamzakhanmarwat@gmail.com}} $|$ 
    \href{https://www.linkedin.com/in/muhammad-hamza-khan-4ab661297/}{\underline{linkedin.com/in/muhammad-hamza-khan}} $|$
    \href{https://github.com/Hamzachan69}{\underline{github.com/Hamzachan69}}
\end{center}

%-----------EDUCATION-----------
\section{Education}
  \resumeSubHeadingListStart
    \resumeSubheading
      {FAST NUCES}{\textbf{Peshawar}}
      {Bachelor of Science in Computer Science,}{Aug. 2022 -- June 2026}
  \resumeSubHeadingListEnd

%-----------EXPERIENCE-----------
\section{Experience}
  \resumeSubHeadingListStart

    \resumeSubheading
      {Unity Intern at Mlabs (MindStorms Studios)}{June 2024 -- Aug 2024}
      {MINDSTORM STUDIOS - Lahore}{}
      \resumeItemListStart
        \resumeItem{Built a solid foundation in the Unity game engine—mastering C\# scripting, physics simulation, asset integration, and core gameplay mechanics to bring interactive experiences to life.}
        \resumeItem{Applied key concepts of game loops, user interaction, and playtesting to refine mechanics and ensure engaging, polished gameplay.}
        \resumeItem{Partnered with a team of four at the Summer Rookie Game Jam ’24 to conceive, prototype, and deliver a 2D slingshot game—handling level design, physics tuning, and iteration under a tight deadline.}
      \resumeItemListEnd

    \resumeSubheading
      {Junior Unity Game Developer at Cognitive Force.}{Nov. 2024 -- Present}
      {COGNITIVE FORCE - Lahore}{}
      \resumeItemListStart
        \resumeItem{Implemented and configured authentication APIs to enable secure user login, registration, and session management.}
        \resumeItem{Developed and integrated user interface components, improving usability and delivering a more intuitive, visually appealing experience.}
        \resumeItem{Assisted in core engine tuning and refactoring, reducing resource consumption and boosting overall application performance.}
      \resumeItemListEnd

    \resumeSubheading
      {Professional Video Editor}{June 2023 -- Present}
      {Upwork.com}{}
      \resumeItemListStart
        \resumeItem{Edited numerous client videos on AI topics into broadcast‑quality content.}
        \resumeItem{Utilized transitions, animated captions, color grading, motion graphics, and visual effects to enhance storytelling and engagement.}
        \resumeItem{Consistently met tight deadlines, delivering on time and fostering high client satisfaction and repeat engagements.}
      \resumeItemListEnd

  \resumeSubHeadingListEnd

%-----------PROJECTS-----------
\section{Projects}
  \resumeSubHeadingListStart

    \resumeProjectHeading
      {\textbf{Snake Game} $|$ \emph{C++ , Visual Studio , Canva , SFML , GitHub}}{March 2023 -- April 2023}
      \resumeItemListStart
        \resumeItem{Remade the classic Nokia Snake game with a nostalgic aesthetic, smoother controls, and polished visuals.}
        \resumeItem{Applied OOP principles (encapsulation, inheritance, abstraction) to structure snake, food, and game manager for clean, maintainable code.}
        \resumeItem{Implemented dynamic gameplay mechanics and custom graphics, optimizing frame updates and collision logic for a responsive, immersive experience.}
      \resumeItemListEnd

    \resumeProjectHeading
      {\textbf{2D SlingShot Game} $|$ \emph{C\# , Unity Game Engine , Visual Studio}}{June 2024 -- Aug 2024}
      \resumeItemListStart
        \resumeItem{Applied internship learnings to architect and develop core gameplay systems during a time‑boxed Game Jam project.}
        \resumeItem{Designed and implemented a physics‑based slingshot mechanic for launching projectiles at zombie targets through destructible obstacles.}
      \resumeItemListEnd

  \resumeSubHeadingListEnd

%-----------TECHNICAL SKILLS-----------
\section{Technical Skills}
 \begin{itemize}[leftmargin=0.15in, label={}]
    \small{\item{
     \textbf{Languages}{\textit{: Python , C / C++ / C\# , SQL} } \\
     \textbf{Developer Tools}{\textit{: Git , SFML , Oracle SQL Developer , Unity Game Engine}.} \\
     \textbf{Databases}{:  \textit{Oracle SQL.} }
    }}
 \end{itemize}

\end{document}
